Getting and cleaning data project
========================================================

Required packages
```{r}
library(reshape2)
library(plyr)
```


```{r}

```

import the features list, remove the hyphens (-) and the brackets from the names and convert to a more "Camel" type variable namingin conventions
```{r}
features <- read.table("features.txt")
#remove brackets
features[,2] <- gsub("\\(\\)","",features[,2])

# Remove commas which are surrounded by numbers
commaSub <- grepl("[0-9],[0-9]",features[,2])
features[commaSub,2] <- gsub(",","to",features[commaSub,2])

#Remove remaining commas
features[,2] <- gsub(",","",features[,2])

#Remove Hyphens and make sure the first letter of the next word is upper case

#function totally half inched from Tyler Rinker on stackoverflow http://stackoverflow.com/questions/11672050/how-to-convert-not-camel-case-to-camelcase-in-r#_=_

camel <- function(x){ #function for camel case
    capit <- function(x) paste0(toupper(substring(x, 1, 1)), substring(x, 2, nchar(x)))
    cam <- function(x) paste(capit(x), collapse="")
    t2 <- strsplit(x, "-")
    sapply(t2, cam)
}

features[,2] <-camel(features[,2])

#Remove remaining braces
brackets <- grepl("Angle\\(",features[,2])
features[brackets,2] <- paste(sub("Angle\\(", "", features[brackets,2]), "Angles", sep="")
features[,2] <- gsub("\\)","", features[,2])
```

Import the training set
```{r}
setwd("./train")
subject.train <- read.table("subject_train.txt")
x.train <- read.table("X_train.txt")
y.train <- read.table("Y_train.txt")
```


Import the test data from working directory
```{r}
setwd("../")
setwd("./test")
subject.test <- read.table("subject_test.txt")
x.test <- read.table("X_test.txt")
y.test <- read.table("Y_test.txt")
```


```{r}
colnames(x.train) <- features[,2]
colnames(y.train) <- "ActivityType"
colnames(subject.train) <- "SubjectId"
train.set <- cbind(y.train, x.train, subject.train)
names(train.set)
```


Make test dataframe
```{r}
colnames(x.test) <- features[,2]
colnames(y.test) <- "ActivityType"
colnames(subject.test) <- "SubjectId"
test.set <- cbind(y.test, x.test, subject.test)
names(test.set)
```


Combine both dataframes together into 1 data frame and swap the activity names for ones that are descriptive
```{r}
whole.set <- rbind(train.set, test.set)
swap <- match(whole.set[,1], seq(1,6))
activity <- c("Walking", "Walking Upstairs", "Walking Downstairs", "Sitting", "Standing", "Laying")
whole.set[which(!is.na(swap)),1] <- activity[na.omit(swap)]
```

Extract all lines which are related to the mean or Standard deviation
```{r}
## creat a logical vector thet finds all references to mean or standard deviation, the rbind ensures that the activity column and the ID column are also inlcuded
measurements <- c(TRUE,grepl("[Mm]ean|[Ss]([Tt])[Dd]", features[,2]),TRUE)
mean.sd <- whole.set[,measurements]
```

create a dataframe of all the means of all the variables split by activity type and subject
```{r}
mean.sd.melt <- melt(data = mean.sd, id.vars = c("ActivityType", "SubjectId") )
mean.sd <-aggregate(value~., data=mean.sd.melt, mean) 
mean.sd <- dcast(mean.sd, ActivityType + SubjectId ~ variable ,value.var="value")
setwd("../")
write.table(mean.sd, "Summary of mean and Std data.txt", sep = ",")
```

PCA
```{r}
PCA.x.test <-princomp(x.test, cor= TRUE)
plot(PCA.x.test, type= "lines")
summary(PCA.x.test)
```

